---
title: "Hands-on Exercise 1A"
author: "yang yayong"
date: "Aug 24 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  message: false
  freeze: true
---

# Hands-on Exercise 1 Geospatial Data Science with R

## 1 Overview

Geospatial Data Science is a process of importing, wrangling, integrating, and processing geographically referenced data sets. In this hands-on exercise, we will learn how to perform geospatial data science tasks in R using the sf package.

## 1.2 Data Acquisition

Data are key to data analytics including geospatial analytics. Hence, before analysing, we need to assemble the necessary data. In this hands-on exercise, we are required to extract the necessary data sets from the following sources:

1.Master Plan 2014 Subzone Boundary (Web) from data.gov.sg

2.Pre-Schools Location from data.gov.sg

3.Cycling Path from LTADataMall

## 1.3 Getting Started

```{r}
pacman::p_load(sf,tidyverse)
```

## 1.4 Importing Geospatial Data

### 1.4.1 Importing polygon feature data in shapefile format

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

### 1.4.2 Importing polyline feature data in shapefile form

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPathGazette")
```

### 1.4.3 Importing GIS data in kml format

```{r}
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")
```

## 1.5 Checking the Content of A Simple Feature Data Frame

### 1.5.1 Working with st_geometry()

```{r}
st_geometry(mpsz)
```

### 1.5.2 Working with glimpse()

```{r}
glimpse(mpsz)
```

### 1.5.3 Working with head()

```{r}
head(mpsz, n=5)  
```

## 1.6 Plotting the Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

## 1.7 Working with Projection

### 1.7.1 Assigning EPSG code to a simple feature data frame

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_transform(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

### 1.7.2 Transforming the projection of preschool from wgs84 to svy21.

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

## 1.8 Importing and Converting An Aspatial Data

### 1.8.1 Importing the aspatial data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

### 1.8.2 Creating a simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

## 1.9 Geoprocessing with sf package

### 1.9.1 Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

```{r}
sum(buffer_cycling$AREA)
```

### 1.9.2 Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

```{r}
summary(mpsz3414$`PreSch Count`)
```

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## 1.10 Exploratory Data Analysis (EDA)

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
