---
title: "Take home Exercise2"
author: "yang yayong"
date: "Sep 27 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  message: false
  freeze: true
---

# **Discovering impacts of COVID-19 on Thailand tourism economy at the province level using spatial and spatio-temporal statistics**

## The scene

Tourism is one of Thailand’s largest industries, accounting for some 20% of the gross domestic product (GDP). In 2019, Thailand earned 90 billion US\$ from domestic and international tourism, but the COVID-19 pandemic caused revenues to crash to 24 billion US\$ in 2020.

Figure below shows the total revenue receipt from tourism sector from January 2019 until Feb 2023. The figure reveals that the revenue from tourism industry have been recovered gradually since September 2021.

However, it is important to note that the tourism economy of Thailand are not evenly distributed. Figure below reveals that the tourism economy of Thailand are mainly focus on five provinces, namely Bangkok, Phuket, Chiang Mai, Sukhothai and Phetchaburi.

::: panel-tabset
## Objectives

As a curious geospatial analytics green horn, you are interested to discover:

-   if the key indicators of tourism economy of Thailand are independent from space and space and time.

-   If the tourism economy is indeed spatial and spatio-temporal dependent, then, you would like to detect where are the clusters and outliers, and the emerging hot spot/cold spot areas.

## The Task

The specific tasks of this take-home exercise are as follows:

-   Using appropriate function of **sf** and **tidyverse**, preparing the following geospatial data layer:

    -   a study area layer in sf polygon features. It must be at [province level](https://en.wikipedia.org/wiki/Provinces_of_Thailand) (including Bangkok) of Thailand.

    -   a tourism economy indicators layer within the study area in sf polygon features.

    -   a derived tourism economy indicator layer in [**spacetime s3 class of sfdep**](https://sfdep.josiahparry.com/articles/spacetime-s3). Keep the time series at **month and year levels**.

-   Using the extracted data, perform global spatial autocorrelation analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-glsa).

-   Using the extracted data, perform local spatial autocorrelation analysis by using [sfdep methods](https://r4gdsa.netlify.app/chap10.html).

-   Using the extracted data, perform emerging hotspot analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-ehsa).

-   Describe the spatial patterns revealed by the analysis above.
:::

## The Data

For the purpose of this take-home exercise, two data sets shall be used, they are:

-   [Thailand Domestic Tourism Statistics](https://www.kaggle.com/datasets/thaweewatboy/thailand-domestic-tourism-statistics) at Kaggle. You are required to use **version 2** of the data set.

-   [Thailand - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-tha?) at HDX. You are required to use the province boundary data set.

## **1Getting Started**

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse,ggplot2)
```

1.2 Getting the Data Into R Environment

Import csv file into r environment

```{r}
tt1923 <- read_csv("data/rawdata/thailand_domestic_tourism_2019_2023_ver2.csv")
```
Import shapefile into r environment

```{r}
b1 <- st_read(dsn = "data/rawdata", 
                 layer = "tha_admbnda_adm1_rtsd_20220121")
```

A tourism economy indicators layer within the study area in sf polygon features.

```{r}
#tb <- b1 %>%
  #left_join(tt1923, by = c("ADM1_EN" = "province_eng"))
```


```{r}
# 筛选出 region_eng 列中为 NA 的行
na_rows <- tt1923[is.na(tt1923$region_eng), ]

# 打印出包含 NA 值的行
na_rows
```

#As can be seen from the figure, na_rows is a data frame containing 0 rows, which #means that is.na(tt1923_selected$region_eng) did not match any NA values, so no rows #were filtered out. Therefore, it can be confirmed that there are no missing values #​​in the region_eng column.

```{r}
unique_counts <- sapply(tt1923, n_distinct)
print(unique_counts)
```
```{r}
unique(tt1923$region_eng)
```


```{r}
unique(tt1923$variable)
```

-   ratio_tourist_stay - ratio of tourists staying

-   no_tourist_stay - number of tourists staying

-   no_tourist_all - total number of tourists

-   no_tourist_thai - number of local tourists

-   no_tourist_foreign - number of foreign tourists

-   revenue_all - total tourism revenue

-   revenue_thai - revenue from local tourists

-   revenue_foreign - revenue from foreign tourists


```{r}
region_total_revenue <- tt1923 %>%
  filter(variable == "revenue_all") %>%  
  group_by(region_eng) %>%                
  summarise(total_revenue = sum(value, na.rm = TRUE))  
```

```{r}
# 按 total_revenue 降序排列
region_total_revenue <- region_total_revenue %>%
  arrange(desc(total_revenue))

# 绘制条形图
ggplot(region_total_revenue, aes(x = reorder(region_eng, total_revenue), y = total_revenue, fill = region_eng)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Tourism Revenue by Region", x = "Region", y = "Total Revenue") +
  theme_minimal() +
  coord_flip() +
  guides(fill = "none")
```

From the perspective of tourism revenue, Central (Central Region) is the most valuable area because its tourism revenue is significantly higher than other regions, especially Bangkok, as an important transportation hub and cultural center, attracting a large number of domestic and international tourists.


```{r}
tb_sc <- tt1923  %>%
  filter(region_eng == "central") %>%
  select(date, province_eng, region_eng, variable,value)
```

```{r}
central_provinces <- tb_sc %>%
  filter(region_eng == "central") %>%
  select(province_eng) %>%  # 选择省份名称列
  distinct()           # 去重，确保每个省份只出现一次

# 打印中心区域的省份信息
central_provinces
```

```{r}
tb_sc_filtered <- tb_sc %>%
  filter(!(province_eng %in% c("Lopburi", "Chainat")))
```


```{r}
selected_provinces <- c(
  "Bangkok", "Lopburi", "Phra Nakhon Si Ayutthaya", "Saraburi", "Chainat",
  "Nakhon Pathom", "Sing Buri", "Ang Thong", "Nonthaburi", "Pathum Thani",
  "Samut Prakan", "Samut Sakhon", "Ratchaburi", "Kanchanaburi",
  "Samut Songkhram", "Suphan Buri", "Phetchaburi", "Prachuap Khiri Khan"
)
```

```{r}
b1_filtered <- b1 %>%
  filter(ADM1_EN %in% selected_provinces)  %>%
select(ADM1_EN, Shape_Area, Shape_Leng, geometry)
```

```{r}
plot(b1_filtered)
```

```{r}
merged_data <- left_join(b1_filtered, tb_sc_filtered, by = c("ADM1_EN" = "province_eng"))
```

```{r}
# 筛选出 region_eng 列中为 NA 的行
na_rows2 <- merged_data[is.na(merged_data$region_eng), ]

# 打印出包含 NA 值的行
na_rows2
```


From data provided,we can select three economic indicators：

-   Tourist quantity index

```{r}
tourist_numbers_thai <- merged_data %>%
  filter(variable %in% c( "no_tourist_thai"))
```

```{r}
tourist_numbers_foreign <- merged_data %>%
  filter(variable %in% c( "no_tourist_foreign"))
```

```{r}
 # 按年份聚合数据
tourist_numbers_thai_yearly <- tourist_numbers_thai %>%
  mutate(year = lubridate::year(date)) %>%  # 提取年份
  group_by(ADM1_EN, year) %>%  # 按区域和年份分组
  summarise(total_value = sum(value, na.rm = TRUE)) %>%  # 对每年的 value 列求和
  ungroup()
```

```{r}
# 确保 tmap 模式为 plot 模式
tmap_mode("plot")

# 绘制每年一个地图
tm_shape(tourist_numbers_yearly) +
  tm_fill("total_value", 
          style = "jenks", 
          palette = "Blues", 
          title = "Total local Tourist Numbers") +
  tm_facets(by = "year") +  # 使用 year 列，生成每年一个地图
  tm_layout(main.title = "Total local Tourist Numbers",
            main.title.size = 1.2,
            legend.width = 1.2,  # 调整图例宽度
            legend.text.size = 0.8)
```

```{r}
 # 按年份聚合数据
tourist_numbers_foreign_yearly <- tourist_numbers_foreign %>%
  mutate(year = lubridate::year(date)) %>%  # 提取年份
  group_by(ADM1_EN, year) %>%  # 按区域和年份分组
  summarise(total_value = sum(value, na.rm = TRUE)) %>%  # 对每年的 value 列求和
  ungroup()
```

```{r}
# 确保 tmap 模式为 plot 模式
tmap_mode("plot")

# 绘制每年一个地图
tm_shape(tourist_numbers_foreign_yearly) +
  tm_fill("total_value", 
          style = "jenks", 
          palette = "Greens", 
          title = "Total foreign Tourist Numbers") +
  tm_facets(by = "year") +  # 使用 year 列，生成每年一个地图
  tm_layout(main.title = "Total foreign Tourist Numbers",
            main.title.size = 1.2,
            legend.width = 1.2,  # 调整图例宽度
            legend.text.size = 0.8)
```


-   Tourism income index

```{r}
tourist_revenue <- tb_sc_filtered %>%
  filter(variable %in% c( "revenue_thai", "revenue_foreign"))
```

```{r}
ggplot(tourist_revenue, aes(x = variable, y = value, fill = variable)) +
  geom_bar(stat = "identity") +
  labs(title = "Tourism Revenue", x = "Revenue Type", y = "Revenue") +
  theme_minimal()

```

-   Tourist behavior index

```{r}
tourist_behavior <- tt1923_selected %>%
  filter(variable %in% c("ratio_tourist_stay",
                         "no_tourist_stay"))
```









We use R’s mutate() and lubridate packages for date manipulation

Creating a Time Series Cube

```{r}
tt1923_selected2 <- tt1923_selected %>%
  mutate(
    YearMonth = ym(format(date, "%Y-%m"))  # 转换为 Date 格式，表示年月
  )
```


```{r}
# 查看聚合后的数据
print(tb_aggregated_yearmonth)
```



```{r}
tm_shape(b1)+
  tm_fill("region_eng", 
          palette = c("#57bfc0","#7977f3", "#ce77b4","#f67774","#f89974", "#f8d673","#f9f777"),
          title = "Region Name") +
  tm_borders(col = "white")+
  tm_layout(main.title = "Study Area (Thailand)",
            main.title.position = "center",
            main.title.size = 2,
            main.title.fontface = "bold",
            legend.title.size = 1.8,
            legend.text.size = 1.3,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", text.size = 1.5, size = 3, position=c("RIGHT", "TOP")) +
  tm_scale_bar(position=c("LEFT", "BOTTOM"), text.size=1.2) +
  tm_grid(labels.size = 1,alpha =0.2)
```




```{r}
tb_st <-spacetime(tb_selected2, b1,
                      .loc_col = "ADM1_EN",
                      .time_col = "YearMonth")
```

```{r}
is_spacetime_cube(tb_st)
```

```{r}
Gtb_nb <- tb_st %>%
  activate("geometry") %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wt = st_inverse_distance(nb, 
                             geometry, 
                             scale = 1,
                             alpha = 1),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")
```

需要提取年份

```{r}
unique_regions <- unique(tb_selected$region_eng)
print(unique_regions)
```
