---
title: "Take home Exercise2"
author: "yang yayong"
date: "Sep 27 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  message: false
  freeze: true
---

# **Discovering impacts of COVID-19 on Thailand tourism economy at the province level using spatial and spatio-temporal statistics**

## The scene

Tourism is one of Thailand’s largest industries, accounting for some 20% of the gross domestic product (GDP). In 2019, Thailand earned 90 billion US\$ from domestic and international tourism, but the COVID-19 pandemic caused revenues to crash to 24 billion US\$ in 2020.

Figure below shows the total revenue receipt from tourism sector from January 2019 until Feb 2023. The figure reveals that the revenue from tourism industry have been recovered gradually since September 2021.

However, it is important to note that the tourism economy of Thailand are not evenly distributed. Figure below reveals that the tourism economy of Thailand are mainly focus on five provinces, namely Bangkok, Phuket, Chiang Mai, Sukhothai and Phetchaburi.

::: panel-tabset
## Objectives

As a curious geospatial analytics green horn, you are interested to discover:

-   if the key indicators of tourism economy of Thailand are independent from space and space and time.

-   If the tourism economy is indeed spatial and spatio-temporal dependent, then, you would like to detect where are the clusters and outliers, and the emerging hot spot/cold spot areas.

## The Task

The specific tasks of this take-home exercise are as follows:

-   Using appropriate function of **sf** and **tidyverse**, preparing the following geospatial data layer:

    -   a study area layer in sf polygon features. It must be at [province level](https://en.wikipedia.org/wiki/Provinces_of_Thailand) (including Bangkok) of Thailand.

    -   a tourism economy indicators layer within the study area in sf polygon features.

    -   a derived tourism economy indicator layer in [**spacetime s3 class of sfdep**](https://sfdep.josiahparry.com/articles/spacetime-s3). Keep the time series at **month and year levels**.

-   Using the extracted data, perform global spatial autocorrelation analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-glsa).

-   Using the extracted data, perform local spatial autocorrelation analysis by using [sfdep methods](https://r4gdsa.netlify.app/chap10.html).

-   Using the extracted data, perform emerging hotspot analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-ehsa).

-   Describe the spatial patterns revealed by the analysis above.
:::

## The Data

For the purpose of this take-home exercise, two data sets shall be used, they are:

-   [Thailand Domestic Tourism Statistics](https://www.kaggle.com/datasets/thaweewatboy/thailand-domestic-tourism-statistics) at Kaggle. You are required to use **version 2** of the data set.

-   [Thailand - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-tha?) at HDX. You are required to use the province boundary data set.

## **1Getting Started**

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse,ggplot2)
```

1.2 Getting the Data Into R Environment

Import csv file into r environment

```{r}
tt1923 <- read_csv("data/rawdata/thailand_domestic_tourism_2019_2023_ver2.csv")
```

Import shapefile into r environment

```{r}
b1 <- st_read(dsn = "data/rawdata", 
                 layer = "tha_admbnda_adm1_rtsd_20220121")
```

A tourism economy indicators layer within the study area in sf polygon features.

```{r}
tb <- b1 %>%
  left_join(tt1923, by = c("ADM1_EN" = "province_eng"))
```

```{r}
library(ggplot2)
ggplot() +
  geom_sf(data = b1, color = "black", size = 0.5)
```

```{r}
colnames(tb)
```

To minimize memory usage, we can remove the columns that are unnecessary for this study and retain only the relevant ones.

```{r}
tb_selected <- tb %>%
  select(Shape_Leng, Shape_Area, ADM1_EN, date.y, region_eng, variable, value, geometry)
```

```{r}
tb_selected
```
```{r}
# 筛选出 region_eng 列中为 NA 的行
na_rows <- tb_selected[is.na(tb_selected$region_eng), ]

# 打印出包含 NA 值的行
print(na_rows)
```

```{r}
# 删除 tb_selected 中 region_eng 为 NA 的行，并去掉重复
tb_selected2 <- tb_selected %>%
  filter(!is.na(region_eng))   # 筛选出 region_eng 为 NA 的行
```


We use R’s mutate() and lubridate packages for date manipulation

Creating a Time Series Cube

```{r}
tb_selected2 <- tb_selected2 %>%
  mutate(
    YearMonth = ym(format(date.y, "%Y-%m"))  # 转换为 Date 格式，表示年月
  )
```

```{r}
# 对数据进行按年-月聚合，计算每个月 ratio_tourist_stay 的均值
tb_aggregated_yearmonth <- tb_selected %>%
  mutate(YearMonth = format(as.Date(date.y), "%Y-%m")) %>%  # 提取年-月信息
  group_by(ADM1_EN, YearMonth) %>%  # 按地区和年-月进行分组
  summarise(mean_ratio_tourist_stay = mean(value, na.rm = TRUE), .groups = 'drop')  # 计算均值
```

```{r}
# 查看聚合后的数据
print(tb_aggregated_yearmonth)
```




```{r}
# 假设 tb_selected2 包含 County 和 YearMonth 列
tb_complete <- tb_selected2 %>%
  complete(ADM1_EN, YearMonth)  # 完整化时间和空间的组合

# 填充缺失值（如果需要，可以根据具体情况填写其他默认值）
tb_complete <- tb_complete %>%
  mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

# 检查数据
head(tb_complete)
```


```{r}

```

```{r}
tb_selected2 <- tb_selected2 %>%
  group_by(County, YearMonth) %>%
  summarise(across(everything(), mean, na.rm = TRUE))
```



```{r}
tb_st <-spacetime(tb_selected2, b1,
                      .loc_col = "ADM1_EN",
                      .time_col = "YearMonth")
```

```{r}
is_spacetime_cube(tb_st)
```

```{r}
Gtb_nb <- tb_st %>%
  activate("geometry") %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wt = st_inverse_distance(nb, 
                             geometry, 
                             scale = 1,
                             alpha = 1),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")
```


```{r}
tm_shape(b1)+
  tm_fill("region_eng", 
          palette = c("#57bfc0","#7977f3", "#ce77b4","#f67774","#f89974", "#f8d673","#f9f777"),
          title = "Province Name") +
  tm_borders(col = "white")+
  tm_layout(main.title = "Study Area (Thailand)",
            main.title.position = "center",
            main.title.size = 2,
            main.title.fontface = "bold",
            legend.title.size = 1.8,
            legend.text.size = 1.3,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", text.size = 1.5, size = 3, position=c("RIGHT", "TOP")) +
  tm_scale_bar(position=c("LEFT", "BOTTOM"), text.size=1.2) +
  tm_grid(labels.size = 1,alpha =0.2)
```

需要提取年份

```{r}
unique_regions <- unique(tb_selected$region_eng)
print(unique_regions)
```


