---
title: "Take home Exercise2"
author: "yang yayong"
date: "Sep 27 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  message: false
  freeze: true
---

# **Discovering impacts of COVID-19 on Thailand tourism economy at the province level using spatial and spatio-temporal statistics**

## The scene

Tourism is one of Thailand’s largest industries, accounting for some 20% of the gross domestic product (GDP). In 2019, Thailand earned 90 billion US\$ from domestic and international tourism, but the COVID-19 pandemic caused revenues to crash to 24 billion US\$ in 2020.

Figure below shows the total revenue receipt from tourism sector from January 2019 until Feb 2023. The figure reveals that the revenue from tourism industry have been recovered gradually since September 2021.

However, it is important to note that the tourism economy of Thailand are not evenly distributed. Figure below reveals that the tourism economy of Thailand are mainly focus on five provinces, namely Bangkok, Phuket, Chiang Mai, Sukhothai and Phetchaburi.

::: panel-tabset
## Objectives

As a curious geospatial analytics green horn, you are interested to discover:

-   if the key indicators of tourism economy of Thailand are independent from space and space and time.

-   If the tourism economy is indeed spatial and spatio-temporal dependent, then, you would like to detect where are the clusters and outliers, and the emerging hot spot/cold spot areas.

## The Task

The specific tasks of this take-home exercise are as follows:

-   Using appropriate function of **sf** and **tidyverse**, preparing the following geospatial data layer:

    -   a study area layer in sf polygon features. It must be at [province level](https://en.wikipedia.org/wiki/Provinces_of_Thailand) (including Bangkok) of Thailand.

    -   a tourism economy indicators layer within the study area in sf polygon features.

    -   a derived tourism economy indicator layer in [**spacetime s3 class of sfdep**](https://sfdep.josiahparry.com/articles/spacetime-s3). Keep the time series at **month and year levels**.

-   Using the extracted data, perform global spatial autocorrelation analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-glsa).

-   Using the extracted data, perform local spatial autocorrelation analysis by using [sfdep methods](https://r4gdsa.netlify.app/chap10.html).

-   Using the extracted data, perform emerging hotspot analysis by using [sfdep methods](https://is415-gaa-tskam.netlify.app/in-class_ex/in-class_ex05/in-class_ex05-ehsa).

-   Describe the spatial patterns revealed by the analysis above.
:::

## The Data

For the purpose of this take-home exercise, two data sets shall be used, they are:

-   [Thailand Domestic Tourism Statistics](https://www.kaggle.com/datasets/thaweewatboy/thailand-domestic-tourism-statistics) at Kaggle. You are required to use **version 2** of the data set.

-   [Thailand - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-tha?) at HDX. You are required to use the province boundary data set.

## **1Getting Started**

```{r}
pacman::p_load(sf, sfdep,spdep, tmap, tidyverse,ggplot2,plotly,Kendall)
```

1.2 Getting the Data Into R Environment

Import csv file into r environment

```{r}
tt1923 <- read_csv("data/rawdata/thailand_domestic_tourism_2019_2023_ver2.csv")
```
```{r}
#write_rds(tt1923,"data/rds/tt1923.rds")

#tt1923 <- read_rds("data/rds/tt1923.rds")
```

Import shapefile into r environment

```{r}
b1 <- st_read(dsn = "data/rawdata", 
                 layer = "tha_admbnda_adm1_rtsd_20220121")
```

A tourism economy indicators layer within the study area in sf polygon features.

```{r}
na_rows <- tt1923[is.na(tt1923$region_eng), ]

na_rows
```

#As can be seen from the figure, na_rows is a data frame containing 0 rows, which #means that is.na(tt1923_selected$region_eng) did not match any NA values, so no rows #were filtered out. Therefore, it can be confirmed that there are no missing values #​​in the region_eng column.

```{r}
unique_counts <- sapply(tt1923, n_distinct)
print(unique_counts)
```

```{r}
unique(tt1923$region_eng)
```


```{r}
unique(tt1923$variable)
```

-   ratio_tourist_stay - ratio of tourists staying

-   no_tourist_stay - number of tourists staying

-   no_tourist_all - total number of tourists

-   no_tourist_thai - number of local tourists

-   no_tourist_foreign - number of foreign tourists

-   revenue_all - total tourism revenue

-   revenue_thai - revenue from local tourists

-   revenue_foreign - revenue from foreign tourists


```{r}
region_total_revenue <- tt1923 %>%
  filter(variable == "revenue_all") %>%  
  group_by(region_eng) %>%                
  summarise(total_revenue = sum(value, na.rm = TRUE))  
```

```{r}
region_total_revenue <- region_total_revenue %>%
  arrange(desc(total_revenue))

ggplot(region_total_revenue, aes(x = reorder(region_eng, total_revenue), y = total_revenue, fill = region_eng)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Tourism Revenue by Region", x = "Region", y = "Total Revenue") +
  theme_minimal() +
  coord_flip() +
  guides(fill = "none")
```
From the perspective of tourism revenue, Central (Central Region) is the most valuable area because its tourism revenue is significantly higher than other regions, especially Bangkok, as an important transportation hub and cultural center, attracting a large number of domestic and international tourists.


```{r}
tb_sc <- tt1923  %>%
  filter(region_eng == "central") %>%
  select(date, province_eng, region_eng, variable,value)
```

```{r}
central_provinces <- tb_sc %>%
  filter(region_eng == "central") %>%
  select(province_eng) %>%  
  distinct()          

# 打印中心区域的省份信息
central_provinces
```

```{r}
tb_sc_filtered <- tb_sc %>%
  filter(!(province_eng %in% c("Lopburi", "Chainat")))
```


```{r}
selected_provinces <- c(
  "Bangkok", "Lopburi", "Phra Nakhon Si Ayutthaya", "Saraburi", "Chainat",
  "Nakhon Pathom", "Sing Buri", "Ang Thong", "Nonthaburi", "Pathum Thani",
  "Samut Prakan", "Samut Sakhon", "Ratchaburi", "Kanchanaburi",
  "Samut Songkhram", "Suphan Buri", "Phetchaburi", "Prachuap Khiri Khan"
)
```

```{r}
b1_filtered <- b1 %>%
  filter(ADM1_EN %in% selected_provinces)  %>%
select(ADM1_EN, Shape_Area, Shape_Leng, geometry)
```

```{r}
plot(b1_filtered)
```

```{r}
merged_data <- left_join(b1_filtered, tb_sc_filtered, by = c("ADM1_EN" = "province_eng"))
```

```{r}
merged_data_date <- merged_data %>%
  mutate(
    YearMonth = as.integer(format(date, "%Y%m"))
  )
```

```{r}
na_rows2 <- merged_data_date[is.na(merged_data_date$region_eng), ]

na_rows2
```

```{r}
central_provinces2 <- unique(merged_data_date %>% dplyr::pull(ADM1_EN))

print(central_provinces2)
```

From data provided,we can select three economic indicators：

-   Tourist quantity index

```{r}
tourist_numbers_thai <- merged_data_date %>%
  filter(variable %in% c( "no_tourist_thai"))
```

```{r}
tourist_numbers_foreign <- merged_data_date %>%
  filter(variable %in% c( "no_tourist_foreign"))
```


```{r, cache=TRUE}
tourist_numbers_thai_monthly <- tourist_numbers_thai %>%
  mutate(
    year = year(date),
    month = month(date)
  ) %>%
  group_by(ADM1_EN, year, month) %>%  # 按省份、年和月分组
  summarise(total_value = sum(value, na.rm = TRUE)) %>%
  ungroup()
```


```{r, cache=TRUE}
tourist_numbers_thai_yearly <- tourist_numbers_thai %>%
  mutate(year = lubridate::year(date)) %>%  
  group_by(ADM1_EN, year) %>%  
  summarise(total_value = sum(value, na.rm = TRUE)) %>%  
  ungroup()
```

```{r}
tmap_mode("plot")

tm_shape(tourist_numbers_yearly) +
  tm_fill("total_value", 
          style = "jenks", 
          palette = "Blues", 
          title = "Total local Tourist Numbers") +
  tm_facets(by = "year") +  
  tm_layout(main.title = "Total local Tourist Numbers",
            main.title.size = 1.2,
            legend.width = 1.2, 
            legend.text.size = 0.8)
```

```{r, cache=TRUE}
 # 按年份聚合数据
tourist_numbers_foreign_yearly <- tourist_numbers_foreign %>%
  mutate(year = lubridate::year(date)) %>%  
  group_by(ADM1_EN, year) %>%  
  summarise(total_value = sum(value, na.rm = TRUE)) %>%  
  ungroup()
```

```{r}
# 确保 tmap 模式为 plot 模式
tmap_mode("plot")

# 绘制每年一个地图
tm_shape(tourist_numbers_foreign_yearly) +
  tm_fill("total_value", 
          style = "jenks", 
          palette = "Greens", 
          title = "Total foreign Tourist Numbers") +
  tm_facets(by = "year") + 
  tm_layout(main.title = "Total foreign Tourist Numbers",
            main.title.size = 1.2,
            legend.width = 1.2,  # 调整图例宽度
            legend.text.size = 0.8)
```

-   Tourism income index

```{r}
tourist_revenue_thai <- merged_data_date %>%
  filter(variable %in% c( "revenue_thai"))
```

```{r}
tourist_revenue_foreign <- merged_data_date %>%
  filter(variable %in% c( "revenue_foreign"))
```

```{r, cache=TRUE}
 # 按年份聚合数据
tourist_revenue_thai_yearly <- tourist_revenue_thai %>%
  mutate(year = lubridate::year(date)) %>%  
  group_by(ADM1_EN, year) %>%  
  summarise(total_value = sum(value, na.rm = TRUE)) %>%  
  ungroup()
```

```{r}

tmap_mode("plot")


tm_shape(tourist_revenue_thai_yearly) +
  tm_fill("total_value", 
          style = "jenks", 
          palette = "YlOrBr", 
          title = "Total revenue from local Tourist ") +
  tm_facets(by = "year") +  
  tm_layout(main.title = "Total revenue from local Tourist",
            main.title.size = 1.2,
            legend.width = 1.2, 
            legend.text.size = 0.8)
```


```{r, cache=TRUE}
 # 按年份聚合数据
tourist_revenue_foreign_yearly <- tourist_revenue_foreign %>%
  mutate(year = lubridate::year(date)) %>%  
  group_by(ADM1_EN, year) %>%  
  summarise(total_value = sum(value, na.rm = TRUE)) %>% 
  ungroup()
```

```{r}
tmap_mode("plot")


tm_shape(tourist_revenue_foreign_yearly) +
  tm_fill("total_value", 
          style = "jenks", 
          palette = "Purples", 
          title = "Total revenue from foreign Tourist ") +
  tm_facets(by = "year") + 
  tm_layout(main.title = "Total revenue from foreign Tourist",
            main.title.size = 1.2,
            legend.width = 1.2,  
            legend.text.size = 0.8)
```

-   Tourist behavior index

```{r}
ratio_tourist_stay <- merged_data_date %>%
  filter(variable %in% c( "ratio_tourist_stay"))
```

```{r, cache=TRUE}

ratio_tourist_stay_yearly <- ratio_tourist_stay %>%
  mutate(year = year(date)) %>%  
  group_by(ADM1_EN, year) %>%  
  summarise(average_ratio = mean(value, na.rm = TRUE)) %>%  
  ungroup()
```

```{r}
tmap_mode("plot")

tm_shape(ratio_tourist_stay_yearly) +
  tm_fill("average_ratio", 
          style = "jenks", 
          palette = "GnBu", 
          title = "Total Number of Tourist stay ") +
  tm_facets(by = "year") +  
  tm_layout(main.title = "Total Number of Tourist stay",
            main.title.size = 1.2,
            legend.width = 1.2,  
            legend.text.size = 0.8)

```

```{r}
no_tourist_stay <- merged_data_date %>%
  filter(variable %in% c( "no_tourist_stay"))
```

```{r}
 # 按年份聚合数据
no_tourist_stay_yearly <- no_tourist_stay %>%
  mutate(year = lubridate::year(date)) %>%  
  group_by(ADM1_EN, year) %>% 
  summarise(total_value = sum(value, na.rm = TRUE)) %>% 
  ungroup()
```

```{r}
# 确保 tmap 模式为 plot 模式
tmap_mode("plot")

# 绘制每年一个地图
tm_shape(no_tourist_stay_yearly) +
  tm_fill("total_value", 
          style = "jenks", 
          palette = "Spectral", 
          title = "Total Number of Tourist stay ") +
  tm_facets(by = "year") +  
  tm_layout(main.title = "Total Number of Tourist stay",
            main.title.size = 1.2,
            legend.width = 1.2,  
            legend.text.size = 0.8)
```


We use R’s mutate() and lubridate packages for date manipulation

Creating a Time Series Cube

There are four important aspects to the spacetime class:

the data: a data.frame object
the geometry: an sf object
the location identifiers
the time column


```{r, cache=TRUE}
tourist_numbers_thai_yearly_st <- as_spacetime(tourist_numbers_thai_yearly, "ADM1_EN", "year")
```

```{r}
is_spacetime_cube(tourist_numbers_thai_yearly_st)
```


Emerging Hot Spots Analysis (EHSA)


```{r}
# 确保 tourist_numbers_thai_st 是一个 sf 对象，并且激活几何
tourist_numbers_thai_yearly_st2 <- tourist_numbers_thai_yearly_st %>%
  activate("geometry")

# Step 1: 计算几何中心点并设置邻接矩阵和权重
tourist_numbers_thai_yearly_st_nb <- tourist_numbers_thai_yearly_st2 %>%
  mutate(
    # 使用 st_centroid 计算几何中心点
    centroid = st_centroid(geometry),
    nb = include_self(st_contiguity(geometry)),  # 计算邻接关系，包含自身
    wt = st_inverse_distance(nb, centroid, scale = 1, alpha = 1),  # 使用几何中心点进行逆距离加权
    .before = 1
  ) %>%
  set_nbs("nb") %>%
  set_wts("wt")
```

```{r}
gi_stars_tourist_numbers_thai_yearly <- tourist_numbers_thai_yearly_st_nb %>% 
  group_by(year) %>% 
  mutate(
    gi_star = local_gstar_perm(total_value,nb,wt)
  ) %>% 
  tidyr::unnest(gi_star)
```

Mann-Kendall Test
A monotonic series or function is one that only increases (or decreases) and never changes direction. So long as the function either stays flat or continues to increase, it is monotonic.

H0: No monotonic trend

H1: Monotonic trend is present

Mann-Kendall Test on Gi
With these Gi* measures we can then evaluate each location for a trend using the Mann-Kendall test. The code chunk below uses Bangkok.

```{r}
bkg <- gi_stars_tourist_numbers_thai_yearly %>% 
  ungroup() %>% 
  filter(ADM1_EN == "Bangkok") %>% 
  select(ADM1_EN, year, gi_star)
```


```{r}
ggplot(data = bkg, 
       aes(x = year, 
           y = gi_star)) +
  geom_line() +
  theme_light()

```

```{r}
p <- ggplot(data = bkg, 
       aes(x = year, 
           y = gi_star)) +
  geom_line() +
  theme_light()

ggplotly(p)
```

Printing Mann-Kendall test report

```{r}
bkg %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>% 
  tidyr::unnest_wider(mk)
```
n the above result, sl is the p-value < 0.05. With reference to the results, we will reject the hypothesis null and infer that a  downward trend.

Mann-Kendall test data.frame
We can replicate this for each location by using group_by() of dplyr package.

```{r}
ehsa_tourist_numbers_thai_yearly <- gi_stars_tourist_numbers_thai_yearly %>%
  group_by(ADM1_EN) %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>%
  tidyr::unnest_wider(mk)
head(ehsa_tourist_numbers_thai_yearly)
```

We can also sort to show significant emerging hot/cold spots

```{r}
emerging <- ehsa_tourist_numbers_thai_yearly %>% 
  arrange(sl, abs(tau)) %>% 
  slice(1:10)
head(emerging)
```
Performing Emerging Hotspot Analysis

```{r}
ehsa_tourist_numbers_thai_yearly <- emerging_hotspot_analysis(
  x = tourist_numbers_thai_yearly_st_nb, 
  .var = "total_value", 
  k = 1, 
  nsim = 99,
  nb_col = "nb",  # 指定邻接矩阵的列名
  wt_col = "wt"        # 指定权重的列名
)
```

```{r}
ggplot(data = ehsa_tourist_numbers_thai_yearly,
       aes(x = classification)) +
  geom_bar()
```
No Pattern Detected: Most areas (12.5) show no significant spatial hot or cold spot patterns. This shows that in these areas, the data does not show a specific hot or cold spot trend, and may be areas where changes are relatively uniform.

Sporadic Coldspot: There are some areas (around 1-2) that exhibit an intermittent coldspot pattern, which means that these areas exhibit lower values ​​during certain periods of time that are not as good as their surroundings. Significantly.

Sporadic Hotspot: There are a few areas (approximately 1-2) that exhibit an intermittent hotspot pattern, meaning that these areas exhibit higher values ​​during certain periods of time, showing contrast with their surroundings.

Visualising EHSA

```{r}
b1_filtered_ehsa <- b1_filtered %>%
  left_join(ehsa_tourist_numbers_thai_yearly,
            by = join_by(ADM1_EN == location))

```

```{r}
b1_filtered_sig <- b1_filtered_ehsa %>%
  filter(p_value > 0.5)  

tmap_mode("plot")

tm_shape(b1_filtered_ehsa) +
  tm_polygons("classification") + 
  tm_borders(alpha = 0.5) +
  tm_shape(b1_filtered_sig) +
  tm_borders(alpha = 0.4)
```


