---
title: "In-class_EX10"
author: "yang yayong"
date: "Nov 4 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  message: false
  freeze: true
---

# Working with Open Government Data

```{r}
pacman::p_load(tidyverse, sf, tmap, httr, performance)
```

## Importing data

The code chunk below imports multiple csv files in a specified folder and append them into a single tibble data frame.

```{r}
#| eval: false
folder_path <- "data/aspatial"
file_list <- list.files(path = folder_path, 
                        pattern = "^realis.*\\.csv$", 
                        full.names = TRUE)

realis_data <- file_list %>%
  map_dfr(read_csv)
```

## Wrangling data

Write a code chunk to perform the followings: - converting values in *Sale Date* field from character to numerical date format, and - extracting resale and condominium transaction records.

```{r}
#| eval: false
condo_resale <- realis_data %>%
  mutate(`Sale Date` = dmy(`Sale Date`)) %>%
  filter(`Type of Sale` == "Resale" &
           `Property Type` == "Condominium")
```

## Geocoding

```{r}
#| eval: false
postcode <- unique(condo_resale$`Postal Code`)
```

```{r}
#| eval: false
url <- "https://onemap.gov.sg/api/common/elastic/search"
found <- data.frame()
not_found <- data.frame()

for (postcode in postcode){
  query <- list('searchVal'=postcode, 'returnGeom'='Y', 
                'getAddrDetails'='Y', 'pageNum'='1')
  res <- GET(url, query=query)
  if ((content(res)$found)!=0){
    found <- rbind(found, data.frame(content(res))[4:13])
  } else {not_found = data.frame(postcode)
  }
}

```

Tidying field names

```{r}
#| eval: false
found <- found %>%
  select(c(6:8)) %>%
  rename(POSTAL = `results.POSTAL`,
         XCOORD = `results.X`,
         YCOORD = `results.Y`)
```

## Converting to Point Feature Data Frame

-   Write a code chunk to join *condo_resale* and *found*. Name the output *condo_resale_geocoded*.

-   Write a code chunk to convert *condo_resale_geocoded* from tibble data frame to sf point feature data frame.

Join the table

```{r}
#| eval: false
condo_resale_geocoded = left_join(
  condo_resale, found, 
  by = c('Postal Code' = 'POSTAL'))
```

Converting to sf

```{r}
#| eval: false
condo_resale_sf <- st_as_sf(condo_resale_geocoded, 
                            coords = c("XCOORD",
                                       "YCOORD"),
                            crs=3414)
```

```{r}
#| eval: false
postcode <- unique(condo_resale$'postal Code')
```

One asset will two schools and we need to calculate the area of each polygon

```{r}
#| eval: false
sf_polygon <- prov_sf %>%
  st_cast("POLYGON") %>%
  mutate(area= st_area(.))
```

Pick up the biggest province

```{r}
#| eval: false
prov_clean
```

To plot to see whether there is a hole in the map
