---
title: "In-class EX05"
author: "yang yayong"
date: "Sep 23 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  message: false
  freeze: true
---

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

```{r}
hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

```{r}
##| eval: false
wm_q<- hunan_GDPPC %>%
  mutate(nb=st_contiguity(geometry),
         wt=st_weights(nb,
                       style = "w"),
         .before=1)
```


mutate:create 
nb:neighbour list 
wt:

```{r}
##| eval: false
moranI<- global_moran(km_qGDPPC,
                      wm_q$nb,
                      wm_q$wt) 
glimpse(moranI)
```

p-value and variance value

```{r}
set.seed(1234)
```


```{r}
##| eval: false
global_moran_perm <- (wm_q$GDPPC,
                      wm_q$nb,
                      wm_q$wt, 
                      nsim = 999)

 
```
start with 99
 
test
 
local moran's I

```{r}
#| eval: false
lisa <- wm_q %>%
  mutate(local_moran = local_moran(
    GDPPC,np,nsim=99),
    .before=1)  %>%
  unnest(local_moran)

```

pay attention to 10 local measure

mean use 

another is median  already have hh hl

pysal

p_ii_sim look at this one


compare ii with p_ii_sim

use + - colour to look at the plot


```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("p_li_sim") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "p-value of Local Moran's I", main.title.size = 2)

```

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("Ii") +
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(
    main.title = "Local Moran's I of GDPPC",
    main.title.size = 2
  )


```

visualising lisa map
```{r}

```

use some statistic significant
p_ii <0.5


ll 

ll significant so is cluster

gi*
view(hcsa)

inclass show a new way to do it 


